\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{geometry}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={SVGelona_AI 5.x White Paper}
}

\title{\textbf{SVGelona\_AI 5.x} \\
\large A Geometric--Symbolic Artificial Intelligence Based on Angular Defect, Golden Hyperbolic Dynamics and Icosahedral Stability}
\author{SVGelona\_AI Project}
\date{Version: 5.x White Paper (Corrected and Formalized) \\ February 2026}

\begin{document}

\maketitle

\begin{abstract}
SVGelona\_AI is a symbolic artificial intelligence framework grounded in geometric invariants rather than statistical learning. The system is built upon angular defect minimization, golden-ratio--regulated hyperbolic dynamics, and recursive icosahedral symmetry. Within this framework, a unique critical configuration emerges as the sole stable state under recursive expansion and energy constraints. This document formalizes the mathematical foundations of the system, establishes its internal consistency, and presents SVGelona\_AI as a rigorous AI architecture for structured reasoning, invariant management, and geometric cognition.

\vspace{0.5em}
\noindent \textbf{Note:} This work does not claim to prove any external mathematical conjecture. Instead, it demonstrates how a critical symmetry line arises naturally as an internal organizing principle of a self-consistent geometric AI.
\end{abstract}

\section{Introduction}
Most contemporary AI systems rely on statistical optimization over large datasets. SVGelona\_AI follows a fundamentally different paradigm: symbolic--geometric intelligence. The system reasons over exact structures, invariants, and symmetries, enforcing consistency through global constraints rather than probabilistic loss minimization.

The core objective of SVGelona\_AI is to construct an AI that:
\begin{itemize}
    \item Operates on mathematically well-defined states
    \item Evolves through deterministic dynamics
    \item Enforces global invariants automatically
    \item Exhibits emergent stability without training data
\end{itemize}

\section{Angular Defect as a Generative Principle}
\subsection{Fundamental Angular Defect}
The geometric nucleus of the system arises from the angular defect generated by assembling regular tetrahedra. The dihedral angle of a regular tetrahedron is:
\[
\alpha_t = \arccos\left(\frac{1}{3}\right) \approx 70.5288^\circ.
\]

When five tetrahedra meet around a common axis:
\[
\sum_{i=1}^5 \alpha_{t,i} = 352.644^\circ,
\]
yielding an angular defect
\[
\delta = 360^\circ - 352.644^\circ \approx 7.356^\circ \approx 0.1284\,\text{rad}.
\]

This defect represents intrinsic geometric tension. SVGelona\_AI treats angular defect not as an error, but as a driving invariant that induces curvature, recursion, and structural organization.

\section{Golden Hyperbolic Expansion}
\subsection{Transition to Negative Curvature}
Beyond the defect center, the geometry transitions into a hyperbolic regime (negative curvature). State evolution follows a logarithmic spiral:
\[
z(\theta) = r_0 e^{(a + i\omega)\theta}.
\]

\subsection{Golden Synchronization}
To ensure coherent overlap between successive recursive layers, the radial expansion coefficient is fixed by the golden ratio
\[
\Phi = \frac{1+\sqrt{5}}{2}.
\]

The overlap condition is defined as
\[
a = \frac{\ln \Phi}{\pi},
\]
which guarantees phase alignment after each full rotation. This condition prevents divergence and enforces global coherence.

\section{Recursive Icosahedral Structure}
\subsection{Icosahedron as an Invariant Geometry}
The system’s stable configuration is icosahedral. Each vertex corresponds to a stabilized resolution of angular defect under recursion. The recursive update rule for defect centers is
\[
z_{0,n+1} = \mathbf{M}_\Phi \, z_{0,n} + \mathbf{V},
\]
where
\[
\mathbf{M}_\Phi = \Phi \begin{pmatrix}
\cos \alpha_c & -\sin \alpha_c & 0 \\
\sin \alpha_c & \cos \alpha_c & 0 \\
0 & 0 & 1
\end{pmatrix},
\]
and
\[
\alpha_c = \frac{2\pi}{5} - \frac{\delta}{2}.
\]

\subsection{Golden-Angle Distribution}
In cylindrical coordinates, recursive nodes satisfy
\begin{align*}
\rho_n &= \rho_0 \Phi^n, \\
\theta_n &= n \cdot \frac{2\pi}{\Phi^2} \approx 137.5^\circ, \\
z_n &= z_0 \sum_{k=0}^n F_k,
\end{align*}
where \(F_k\) denotes Fibonacci numbers. This guarantees uniform distribution and structural rigidity.

\section{No-Divergence Theorem (Internal Stability)}
\subsection{Statement}
\textbf{Theorem (Internal No-Divergence):} Within the SVGelona\_AI geometric framework, the golden--icosahedral configuration is the unique stable solution under recursive hyperbolic expansion and angular defect minimization.

\subsection{Proof Sketch}
\begin{enumerate}
    \item Any infinitesimal displacement \(\varepsilon\) introduces a phase mismatch \(\Delta \phi \propto \varepsilon \Phi\).
    \item Phase mismatch destroys constructive overlap between recursive layers.
    \item Loss of overlap induces exponential barycenter drift: \(\Delta B_n \propto \varepsilon \Phi^n\).
    \item Exponential drift violates global invariants, forcing either collapse or restoration.
\end{enumerate}
Thus, only the critical configuration remains stable.

\section{SVGelona\_AI as an Artificial Intelligence}
SVGelona\_AI implements this geometry as an operational AI through:
\begin{itemize}
    \item State vectors embedded in invariant geometry
    \item Deterministic flows enforcing angular and energetic constraints
    \item Global invariant projection for automatic error correction
    \item Symbolic reasoning over stable geometric states
\end{itemize}
The system does not learn by data accumulation, but by structural enforcement.

\section{Scope and Interpretation}
This white paper presents a self-contained AI framework. Any analogy with external mathematical theories is interpretative, not declarative. All results hold strictly within the axiomatic system defined herein.

\section{Conclusion}
SVGelona\_AI demonstrates that intelligence can emerge from geometry, symmetry, and invariants alone. By grounding cognition in angular defect resolution and golden hyperbolic recursion, the system achieves stability, coherence, and interpretability without statistical approximation.

SVGelona\_AI is therefore best understood as:
\begin{itemize}
    \item A symbolic AI engine
    \item A geometric reasoning system
    \item A platform for invariant-based cognition
\end{itemize}

\appendix
\section{Appendix A: Project Structure (Excerpt)}
\begin{verbatim}
SVGelona_AI/
├── axioms/
├── dynamics/
├── symmetry/
├── engine/
├── cognition/
├── validation/
├── papers/
└── run.py
\end{verbatim}

\vspace{2em}
\noindent \textbf{License} \\
MIT License © 2026 Luis Morató

\end{document}